---
title: "STA302 Final Project: Reproducible Daily Analysis"
author: "Haiwen Yi and group members"
date: "2026-09-20"
output: html_document
---

## Purpose

This file is the submission entry point for the complete daily HEWJ–EWJ analysis. It executes the accompanying R script, which reads only the frozen files in `data/raw`, writes the cleaned analysis table, fits the declared models, computes classical and Newey–West HAC(5) inference, performs diagnostics and chronological validation, and regenerates all figures and result tables.

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(file.path(getwd(), "..")))
```

## Reproduce all results

```{r run-all, echo=TRUE, message=TRUE, warning=TRUE}
# Knit from this repository. The script requires only R (>=4.4) and jsonlite.
source("analysis/run_analysis.R")
```

## Inspect primary results

```{r primary-results, echo=TRUE}
read.csv("results/coefficients_classical.csv")
read.csv("results/coefficients_hac5.csv")
read.csv("results/diagnostics.csv")
read.csv("results/chronological_validation.csv")
```

## Figures

```{r figures, echo=TRUE, out.width='100%'}
project_root <- normalizePath(".")
knitr::include_graphics(c(
  file.path(project_root, "figures", "fx_slope_by_period.png"),
  file.path(project_root, "figures", "residual_diagnostics.png"),
  file.path(project_root, "figures", "coefficient_intervals_hac5.png")
))
```
